A Handbook for Tenured Track Faculty in Computer Science in the US

### 2024-05-23 faculty.tex
updates

\chapter{Research}
\chapter{Obtaining Funding}
Getting funding is arguably the most important part of getting tenure.  Most new professors struggle with grant writing, which is very different than writing research papers. However, without funding, you can't pay your students, travel to conferences, or even pay your own salary. In short, the lack of funding cause lost of stress and anxiety, and threaten your tenure.  In this chapter, we will discuss the various ways to obtain funding.

@@

\subsection{Volunteer for more Leadership roles} Once you have served on a few PCs and NSF panels, you should aim for more leadership roles.  For example, you can volunteer to be a chair of certain sessions/tracks of a large conference, or even a general chair of a workshop or conference.  These roles are more time consuming, but they can be more rewarding (e.g., more people will know about you) and in general can help your career, especially your tenure case.

@@

\subsection{Personal Experience} I have served on many PCs and NSF panels. I mostly cold email people for these opportunities. For example, I directly email OOPSLA and PLDI PC chairs and asked, and they accepted my request (I did have papers at OOPSLA and PLDI when I asked). I tried this with SE conferences  e.g., ICSE (and also nominate myself through the official form), and it never works, even though I have many papers at ICSE.
I also emailed NSF program managers and asked to serve on a panel, and they accepted my request.

I felt that I didn't get much out of serving on PC and didn't really like it much. However, I still do it because I feel I should give back to the community.
On the other hand, I thoroughly enjoy serving on NSF panels (though of course my proposals still got rejected frequently). While the work is more intense and takes more time, I found it more rewarding and enjoy the discussions. Since my first NSF panel in 2019, I have served 7 panels (at least once every year since '19) and I look forward to serving more.

@@

\subsection{Personal} Since coming to GMU, I have been very active in serving the department.  I share the vision of the department and want to help achieve it. In addition to the typical services load, I volunteer to serve in additional committees and leadership roles. For example, I maintain the CSRankings DB for the department, organize the Open House to recruit admitted students, participate in the Dept Chair renewal committee, etc.
Even though I am not on the search committee, I meet with every faculty candidate (I talked to 22 people in Spring'23, which was my study break) and help recruit strong ones. I also agree to be director of the MS program, which is a lot of work and typically done by tenured faculty.

Sometimes I feel overwhelmed by the services, but I also feel that I am making a difference in the department and that is rewarding.
I am also building a good relationship with my colleagues (and also head butt with some), and that is important for my tenure case.
In addition, I learn a lot from being in various roles (e.g., how class scheduling work, how the department is run, how to recruit students and recruit faculty, how to run a colloquium, etc) and feel more confident in my leadership skills.

@@

\chapter{Tenure Package and Application}
After 5 or 6 years, it is time for you to go up (or apply) for tenure.  You will need to prepare a tenure package and submit it to your department.  The department will then review your package and vote on whether to recommend you for tenure.  If the department votes in your favor, the package will then go to the college, and then to the university.  The university will then make the final decision on whether to grant you tenure.

This document shows my experience in preparing my tenure package and applying for tenure. I will share my timeline, the package (statements, letter writer sections), and the process.  I will also share some advice on this process.

@@

\begin{itemize}
    \item Fall 2021: Moved to GMU (I started at UNL in 2016)
    \item (late) Fall 2022: I told my department chair that I plan to go up for tenure in 2023.
    \item Spring 2023: I started preparing my tenure package, which includes writing statements and think reference letters.  I also took a semester off because GMU has a pre-tenure study break policy.
    \item Summer 2023: I submitted my tenure package.
    \item Fall 2024: I gave a tenure talk. Department voted in my favor, and then the college voted in my favor, and the decision goes up to the university.
    \item Spring 2024: In March, I got the official notification that I got tenure and promoted to Associate Prof., beginning in Fall 2024.
\end{itemize}

@@

\subsection{Letters}
This is a huge part of your tenure package.  You will need to ask for letters from senior people in your area.  Someone who can vouch for your research, teaching, and services.
You will need 4-6 letters. However, you will not be the one asking for these letters, but your department will. You can suggest who to ask.

Typically, there are two lists of letter writers: one consisting of people suggested by you and the other consisting of people suggested by the department (e.g., people who work in your field in your dept). There can also be another list, the ``blacklist'' where you list people that you don't want to write letters for you.

Originally I asked several senior people to write my letters. Then I found out that there is a rule at GMU that I cannot ask people whom have collaboration with me in the last 7 years!!! That invalidates most of my letter writers.  Moreover, GMU (and likely most other universities) would prefer writers not from my list (because people on my list are likely to be biased), and has a rule that at least 60\% of the letter writers must be from the department's list. Moreover, there's a good chance that a letter writer will decline the invitation, so there should be more names.

@@

>>>>

\chapter{Research}
\chapter{Obtaining Funding}
Getting funding is arguably the most important part of getting tenure.  Most new professors struggle with grant writing, which is very different compared to research papers. However, without funding, you can't pay your students, conference registration and travel, or even pay your own summer salary. In short, the lack of funding causes lost of stress and anxiety, and threaten your tenure.  In this chapter, we will discuss the various ways to obtain funding.

@@

\subsection{Volunteer for more Leadership roles} Once you have served on a few PCs and NSF panels, you should aim for more leadership roles.  For example, you can volunteer to be a chair of certain sessions/tracks of a large conference, or even a general chair of a workshop or conference.  These roles are more time-consuming, but they can be more rewarding (e.g., more people will know about you) and in general can help your career, especially your tenure case.

@@

\subsection{Personal Experience} In my case I mostly cold email people for opportunities. For example, I directly email OOPSLA and PLDI PC chairs and asked, and they accepted my request (I did have papers at OOPSLA and PLDI when I asked). I tried this with SE conferences  e.g., ICSE (and also nominate myself through the official form), and it never works, even though I have many papers at ICSE. I also got invited to serve on a PC because a senior colleague recommended me.
For NSF, I emailed program managers and asked to serve on their panels.

I felt that I didn't get much out of serving on PCs and didn't really like the review process much. However, I still do it because I feel it is my responsibility give back to the community.

On the other hand, I thoroughly enjoy serving on NSF panels (though of course my proposals still got rejected frequently). While the work is more intense and takes more time, I found it more rewarding and enjoy the discussions. Since my first NSF panel in 2019, I have served 8 panels (at least once every year since '19) and I look forward to serving more.

@@

A good department will try to match you with committees that can help junior faculty build their lab and tenure case (and shield you from committees that are not as useful).

\subsection{Personal Experience} Since coming to GMU, I have been very active in serving the department.  I share the vision of the department and want to help achieve it. In addition to the typical services load, I volunteer to serve in additional committees and leadership roles. For example, I maintain the CSRankings DB for the department, organize the Open House to recruit admitted students, participate in the Dept Chair renewal committee, etc.
Even though I am not on the search committee, I meet with every faculty candidate (I talked to 22 people in Spring'23, which was my study break) and help recruit strong ones. I also agree to be director of the MS SWE program, which is a lot of work and typically done by tenured faculty.

Occasionally I feel overwhelmed by having many services, but I also feel that I am making a difference in the department and that is rewarding.
I am also building a good relationship with my colleagues (and also head butt with some), and that is important for my tenure case.
In addition, I learn a lot from being in various roles (e.g., how class scheduling work, how the department is run, how to recruit students and recruit faculty, how to run a colloquium, etc).

@@

\chapter{Renewal and Tenure Application}

Toward the end of your 3rd year, you will need to prepare a renewal package and submit it to your department.  The department will then review your package and vote on whether to renew you for your 4th (and perhaps also 5th year). Your next renewal is essentially your tenure application. The department will vote on whether to recommend your renewal.

Similarly, after 5 or 6 years, it is time for you to go up (or apply) for tenure.  You will need to prepare a tenure package and submit it to your department.  The department will then review your package and vote on whether to recommend you for tenure.  If the department votes in your favor, the package will then go to the college, and then to the university (e.g., provost, president, board of visitors, etc).  The university will then make the final decision on whether to grant you tenure.

%I will focus more on the tenure application process, as the renewal process is similar but less  intense.

%are so important, you should start preparing for them early.  You should also talk to your department chair and mentor about the process and what is expected of you.  You should also talk to your colleagues who have gone through the process and ask for their advice.

%This document shows my experience in preparing my tenure package and applying for tenure. I will share my timeline, the package (statements, letter writer sections), and the process.  I will also share some advice on this process.

@@

\begin{itemize}
    \item Fall 2021: Moved to GMU with accelerated tenure clock (meaning I don't have to restart from the beginning as I have already been at UNL since 2016).
    \item (late) Fall 2022: I told my department chair that I plan to apply for tenure in 2023 (which means if things go well then I get tenured in Fall'24).
    \item Spring 2023: I started preparing my tenure package, which includes writing statements and think reference letters.  I also took a semester off because GMU has a pre-tenure study break policy.
    \item Summer 2023: I submitted my tenure package.
    \item Fall 2024: I gave a tenure talk. Department voted in my favor, and then the college voted in my favor, and the decision goes up to the university.
    \item Spring 2024: In May, I got the official notification that I got tenured and promoted to Associate Prof., beginning in Fall 2024.
\end{itemize}

@@

\subsection{Letters}
This is a huge part of your tenure package.  You will need letters from senior people in your area.  Someone who can vouch for you.
You will need 4--6 letters. However, you will not be the one asking for these letters, but your department will. You can suggest who to ask.

Typically, there are \emph{two lists} of letter writers: one consisting of people suggested by you and the other consisting of people suggested by the department (i.e., your dept will come up with a list of people they think are good letter writers for you). There can also be another list, the ``blacklist'' where you list people that you don't want to write letters for you.

\subsubsection{Personal Experience}
Originally I asked several senior people to write my letters. Then I found out that there is a rule at GMU that I cannot ask people whom have collaboration with me in the last 7 years!!! That invalidates most of my letter writers.  Moreover, GMU (and likely most other universities) would prefer writers not from my list (because people on my list are likely to be biased), and has a rule that at least 60\% of the letter writers must be from the department's list. Moreover, there's a good chance that a letter writer will decline the invitation, so there should be more names.

@@

\paragraph{Tenure Visits} Many faculty will visit other universities to give talks and meet people.  This is a good way to get your name out and to get letters.  For example, if you plan to put person X on your list or you think person X will likely be selected by the department, you can ask X to visit their university and give a talk.

\paragraph{Inviting speakers to colloquium} Many departments have colloquium series where they invite speakers to give talks.  You can invite potential letter writers to your department's colloquium.

In general, these are a good way to get to know people in your field and to get your name out. Many of these, especially tenure visits, are common practices for junior faculty applying for tenure.

\section{Renewal Package}
